A Formal Model of Checked C

### 2023-04-05 sandp23/introduction.tex
Adding
Addiong all

Vulnerabilities due to memory corruption, especially spatial memory corruption, 
are still a major issue for C programs~\cite{cvetrend, microsoftmemsafe, Zeng:2013:SRF:2534766.2534798} 
despite many efforts that tried to prevent them~\cite{song2019sanitizing}.
Several industrial and research efforts, including CCured~\cite{Necula2005},
Softbound~\cite{softbound}, and ASAN~\cite{Serebryany2012},
have investigated ways to better compile C programs with automatic spatial safety enforcement.
These approaches all impose performance overheads deemed too high for deployment use. 
% 
%\liyi{Should we say this? Is the paper about overhead reducing? }
Recently,~\citet{Elliott2018} and~\citet{li22checkedc} introduced and formalized \checkedc, an
open-source extension to C,
to ensure a program’s spatial safety by introducing new pointer types,~\ie checked ($\cmode$) pointer types.
%
The checked pointers are represented as system-level memory words without ``fattening'' metadata~\cite{duck2016heap}, 
and ensuring backward compatibility,~\ie developers can use checked and regular (unchecked $\umode$) pointers within the same program.
%users are able to split code into checked and unchecked regions and incrementally convert
%C code in the unchecked region to Checked C code in the checked one.
However, as we explain in~\sect{subsec:nosafetyagsintuncheckedcode}, the
unconverted or unchecked ($\umode$) code can violate guarantees provided in $\cmode$
regions.
%
We need to ensure that~\emph{code executed as part of 
unchecked ($\umode$) regions does not lead to the safety violations in checked ($\cmode$) regions  with the use of program partitioning mechanism \cite{rul2009towards}}.

@@

We evaluated~\systemname by partioning seven large real-world programs to demonstrate its effectiveness.
Our evaluation shows that~\systemname provides a flexible, low-overhead program partitioning mechanism and guarantees spatial memory safety.

>>>>

Vulnerabilities due to memory corruption, especially spatial memory corruption, 
are still a major issue for C programs~\cite{cvetrend, microsoftmemsafe, Zeng:2013:SRF:2534766.2534798} 
despite many efforts that tried to prevent them~\cite{song2019sanitizing}.
Several industrial and research efforts, including CCured~\cite{Necula2005},
Softbound~\cite{softbound}, and ASAN~\cite{Serebryany2012},
have investigated ways to better compile C programs with automatic safety enforcement.
These approaches impose performance overheads deemed too high for deployment use, especially for resource-constrained embedded systems.
%
Previous works~\cite{meng2021bran, du2019leopard} show that vulnerabilities are not uniformly distributed in real programs.
Some components/functions are likelier to have a certain class of vulnerabilities.
\eg{} spatial safety issues in string processing functions.
have.

Program partitioning is a well-known technique that enables separating potentially vulnerable functions (untrusted region) from the rest of the program (trusted region).
%
Several automated program partitioning techniques~\cite{tan2017principles, brumley2004privtrans, bittau2008wedge, lind2017glamdring, liu2017ptrsplit} exist, but they have a considerable performance overhead.

One of the main reasons for this is combining partitioning with~\acf{SFI}.
Specifically, in addition to partitioning, existing techniques also provide~\ac{SFI} such that code in the untrusted region cannot affect code in the trusted region.\eg{} by executing untrusted partition in a separate process~\cite{liu2017ptrsplit} or executing in a Trusted Execution environment~\cite{lind2017glamdring}.
However, this extreme partitioning is not always necessary and feasible in certain resource constrained environments.
For instance, we want to isolate a string processing function only from spatial violations,~\ie any spatial violations (\eg buffer overread) in that function should not affect the program.

But existing partitioning techniques are specialized for a specific~\ac{SFI}~\cite{wahbe1993efficient} technique.
For instance,~\ptrsplit{}~\cite{liu2017ptrsplit} technique assumes that partitions are hosted as separate processes and data is exchanged through marshaling.
Any other isolation mechanism or data exchange (\eg shared memory) cannot be supported as the partitioning technique is specialized for marshaling by using~\ac{PDG} and tracking bounds information.
In summary, existing techniques combine policy (\ie{}~\emph{what} to partition) with mechanism (\ie{}~\emph{how} to partition).
We argue that existing program partitioning mechanisms are restrictive and inflexible for different use cases (\eg isolation from spatial safety vulnerabilities). We need a technique that enables developers to flexibility partition programs and enforce isolation -- balancing safety guarantees and performance overhead.

% 
%\liyi{Should we say this? Is the paper about overhead reducing? }
Recently,~\citet{Elliott2018} and~\citet{li22checkedc} introduced and formalized \checkedc, an
open-source extension to C,
to ensure a program’s spatial safety by introducing new pointer types,~\ie checked ($\cmode$) pointer types.
%
The checked pointers are represented as system-level memory words without ``fattening'' metadata~\cite{duck2016heap}, 
and ensuring backward compatibility,~\ie developers can use checked and regular (unchecked $\umode$) pointers within the same program.
%users are able to split code into checked and unchecked regions and incrementally convert
%C code in the unchecked region to Checked C code in the checked one.
However, as we explain in~\sect{subsec:nosafetyagsintuncheckedcode}, the
unconverted or unchecked ($\umode$) code can violate guarantees provided in $\cmode$
regions.
%
We need to ensure that~\emph{code executed as part of 
unchecked ($\umode$) regions does not lead to the safety violations in checked ($\cmode$) regions  with the use of program partitioning mechanism \cite{rul2009towards}}.

@@

We evaluated~\systemname~\footnote{Our implementation is available open source at~\oururl{}.} by partioning seven large real-world programs to demonstrate its effectiveness.
Our evaluation shows that~\systemname provides a flexible, low-overhead program partitioning mechanism and guarantees spatial memory safety.
